### 2022-05-31 article.tex
Badges are useful

Here we define paradata as the data that describes the generation of data.
In genetic epidemiology, the data generated is mostly the results 
of an analysis (e.g. predicting a person having a disease),
as done by computer code.
In such context, paradata is usually the scientific paper that
describes what the computer code does.
However, this has the unrealtistic 
assumption that there is a perfect match between the paper and the code.
In this chapter it is argued that the source code should should be supplied,
as this is the true paradata: if the paper and code disagree, it is the
code that has generated the results.
The chapter concludes by some rules how to allow code to serve as paradata,
and hence allowing computational research to be truly reproducible.

@@

{\bf Keywords:} paradata, reproducible research, code, software,
FAIR data, computational research, Open Science, DevOps, best practices,
genetic epidemiology

@@

\begin{table}[h]
  \begin{tabular}{lll}
    Term      & Definition                                            & Example                                                         \\
    \hline
    Code      & Body of text to be run by a computer                  & The scripts run by a computational experiment                   \\
    Data      & Individual facts, statistics, or items of information & A SNP that has a significant association                        \\
    Genotype  & The DNA allele at a certain location                  & AA, AC, CG, GT, ...                                             \\
    Paradata  & Data that describes the generation of data            & The code to conclude that a SNP has a significant association   \\
    Phenotype & How an organism looks like in the broadest sense      & The concentration of IL6RA in the blood                         \\
    Metadata  & Data that provides information about other data       & The article that describes an experiment                        \\
    Trait     & A phenotype                                           & The concentration of IL6RA in the blood                         
  \end{tabular}
\end{table}

@@

Some code hosts, among others, GitHub, 
allows the researcher to start software scripts when uploading a new version.
To determine the reproducibility of the code, one such script should
run the code (on trivial data) to verify if the code works at all.
If the functionality of the code is lost, the researcher
can immediatly see which small change was the cause of this.
A formalized version of this workflow is called continuous integration (CI).
Using a CI service (such as GitHub Actions), is known to significantly 
increase the number of bugs exposed (\cite{vasilescu2015}) and increases
the speed at which new features are added (\cite{vasilescu2015}).

@@

A more sophisticated check for the code's correctness is
to run units tests and determine the percentage of code that has been
tested, which is called the code coverage. Code coverage
correlates with code quality (\cite{horgan1994,del1995correlation}). 
The non-profit organisation rOpenSci, which peer-reviews code,
has made it a prerequisite to have a code coverage of 100\%.

@@

@@

The most reproducible way of submitting the code of an experiment,
is by providing the code and all it needs 
installed in a virtual container.
Virtualization allows one to provide a standard environment,
that can be used on any other computer.
The best known examples of container software are Docker and Singularity,
where the latter is the only viable option to be able to run
an experiment on a high performance computing (HPC) environment,
such as a computer cluster.

@@

@@

\paragraph{Definition}

@@

Publishing code may be disadvantageous for an author.
For science, yes, as this allows reproducible research.
For an author, publishing code alongside an experiment opens up
the possibily to receive questions regarding that code.
For a publisher
For a knowledge archivist,

>>>>

Here we define paradata as the data that describes the generation of data.
In genetic epidemiology, the data generated is mostly the results 
of an analysis (e.g. predicting a person having a disease),
as done by computer code.
In such context, paradata is usually the scientific paper that
describes what the computer code does.
However, this has the unrealtistic 
assumption that there is a perfect match between the paper and the code.
In this chapter it is argued that the source code should should be supplied,
as this is the true paradata: if the paper and code disagree, it is the
code that has generated the results.
The chapter concludes by some rules how to better code to serve as paradata,
and hence allowing computational research to be truly reproducible.

@@

{\bf Keywords:} paradata, reproducible research, code, software,
FAIR data, computational research, Open Science, best practices,
genetic epidemiology

@@

\begin{table}[h]
  \begin{tabular}{lp{5cm}p{5cm}}
    Term      & Definition                                            & Example                                                         \\
    \hline
    Code      & Body of text to be run by a computer                  & The scripts run by a computational experiment                   \\
    Data      & Individual facts, statistics, or items of information & A SNP that has a significant association                        \\
    Genotype  & The DNA allele at a certain location                  & AA, AC, CG, GT, ...                                             \\
    Paradata  & Data that describes the generation of data            & The code to conclude that a SNP has a significant association   \\
    Phenotype & How an organism looks like in the broadest sense      & The concentration of IL6RA in the blood                         \\
    Metadata  & Data that provides information about other data       & The article that describes an experiment                        \\
    Trait     & A phenotype                                           & The concentration of IL6RA in the blood                         
  \end{tabular}
\end{table}

@@

Some code hosts, among others, GitHub, 
allows the researcher to start software scripts when uploading a new version.
To determine the reproducibility of the code, one such script should
run the code (on trivial data) to verify if the code works at all.
If the functionality of the code is lost, the researcher
can immediatly see which small change was the cause of this.
A formalized version of this workflow is called continuous integration (CI).
Using a CI service (such as GitHub Actions), is known to significantly 
increase the number of bugs exposed (\cite{vasilescu2015}) and increases
the speed at which new features are added (\cite{vasilescu2015}).
Using CI opens up the possibility to informally annotate features of
a repository. For example, Figure \ref{fig:build_badge} shows a 
build badge that indicates that a LaTeX document (i.e. this paper)
could be built.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{build_badge.png}
  \caption{
    A build badge for the GitHub repository of this article.
    The description indicates which process it is that the badge
    signals success (as in this case) or failure of. 
  }
  \label{fig:build_badge}
\end{figure}

@@

A more sophisticated check for the code's correctness is
to run units tests and determine the percentage of code that has been
tested, which is called the code coverage. Code coverage
correlates with code quality (\cite{horgan1994,del1995correlation}). 
The non-profit organisation rOpenSci, which peer-reviews code,
has made it a prerequisite to have a code coverage of 100\%.
Also here, CI opens up the possibility to informally annotate the
code coverage. For example, Figure \ref{fig:badge_codecov} shows a 
build badge that indicates that certain R package's code 
has a 100\% code coverage.

\begin{figure}[!htbp]
  \centering
  \includegraphics[]{badge_codecov.png}
  \caption{
    A build badge for the code coverage of an R 
    package (\url{https://github.com/ropensci/beautier}).
    The umbrella signals that the code coverage is hosted by
    CodeCov (\url{https://codecov.io/}). 
    The text indicates which process it is that the badge
    signals, which in this case, is that the code is fully covered
    by tests.
  }
  \label{fig:badge_codecov}
\end{figure}

@@

\paragraph{Suggestions for knowledge management}

Here are multiple suggestions to make hosted code FAIR.
To make the hosted code findable, commonly a paper refers to its URL.
At the homepage of hosted code, there is no standardized metadata, to,
for example, link to the paper.
Using a popular code host ensures the code is easily accessible.
There is some interoperabily supplied by a version control manager,
such as in obtaining the history of the code.
The reusability of code is reasonably easy to assess, by using build badges.
However, there is no guarantee that a build badge that claimes success,
actually is an honest signal (i.e. that code may not actually build),
nor are there standards for what 'success' entails.

@@

The most reproducible way of submitting the code of an experiment,
is by providing the code and all it needs 
installed in a virtual container.
This is close to the golden standard suggested by 
\cite{peng2011reproducible} (see also Figure \ref{fig:peng2011reproducible}),
however, that was before virtualization existed.
Virtualization allows one to provide a standard environment,
that can be used on any other computer.
The best known examples of container software are Docker and Singularity,
where the latter is the only viable option to be able to run
an experiment on a high performance computing (HPC) environment,
such as a computer cluster.

@@

\paragraph{Suggestions for knowledge management}

Here are multiple suggestions to make containers with runnable code FAIR.
To make such containers findable, the most common 
environments (Syslabs and DockerHub) to upload containers must be indexed.
Uploading a container to those environments is, however, limited by
a restricted amount of free storage, with the possible consequence that
those containers are uploaded somewhere where they escape proper indexing.
Having containers uploaded at the existing websites ensures they are
accessible, interoperable and reusable.
One could argue however, that the scientific community need its own,
similar website to upload containers, as the metadata for these
sites are limited to free-text comments and tags (DockerHub)
or not metadata at all (Syslabs).

@@

\paragraph{Definition of paradata}

@@

Publishing code may be disadvantageous for an author.
For science, yes, as this allows reproducible research.
For an author, publishing code alongside an experiment opens up
the possibily to receive questions regarding that code.
Note, however, that not publishing code will always thwart
the reproduction of incorrect code, at the cost of a scientific
career \cite{baggerly2009deriving}.
